---
title: "Anomaly plotting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# devtools::install_github("tomz23/OpenSourceAP.DownloadR")
library(OpenSourceAP.DownloadR)
```

To plot the long-short return of a particular anomaly:
1. Download the portfolio data
2. (Optional) Download the signal documentation data
3. Calculate cumulative returns over time and plot

# Open an instance to the OpenAP dataset:
```{r}
openap <- OpenAP$new()
```

# 1. Dowload anomaly portfolio data
```{r}
port_op = openap$dl_port('op', 'IndIPO')
```


# 2. (Optional) Download signal documentation and extract some info
```{r}
# Download signal doc
signaldoc = openap$dl_signal_doc()

# Extract some info
DateInSampleStart = as.Date(paste0(signaldoc$SampleStartYear[signaldoc$Acronym == 'IndIPO'], '-12-31'))
DateInSampleEnd   = as.Date(paste0(signaldoc$SampleEndYear[signaldoc$Acronym == 'IndIPO'], '-12-31'))
DatePub           = as.Date(paste0(signaldoc$Year[signaldoc$Acronym == 'IndIPO'], '-12-31'))

```

3. Calculate cumulative returns over time and plot
```{r}
# Calculate cumulative returns
cumRet = port_op %>%
  arrange(port, date) %>% 
  group_by(port) %>% 
  mutate(cumret = cumprod(1 + ret/100))

cumRet %>% 
  filter(port == 'LS') %>% 
  ggplot(aes(x = date, y = cumret)) +
  geom_line() +
  geom_vline(xintercept = c(DateInSampleStart, DateInSampleEnd, DatePub), linetype = 2) +
  labs(y = 'Cumulative return', x = 'Date') +
  theme_bw()
# +
#   geom_text(aes(y = max(cumRet$cumret), label = c('In sample end', 'Publication')), vjust = -1) 


```


